Magnetised Polish doughnuts revisited

### 2017-11-14 magnetised_polish_doughnuts_revisited/paper_lang_rev/to_send.tex
new paper added

The investigation of this type of systems, either by analytical or numerical means, may rely on the ability to construct suitable 
representations based on physical assumptions. The construction of equilibrium models of stationary discs around black holes has indeed a long tradition (see~\cite{Abramowicz:2013} and references therein). In particular, the so-called ``Polish doughnuts"~\citep{Abramowicz:1978,Kozlowski:1978} provide a  very general method to build equilibrium configurations of perfect fluid matter orbiting around a Kerr black hole. This fully relativistic model assumes that the disc is non-magnetised and that the matter obeys a  constant specific angular momentum distribution. This method was later extended by~\cite{Komissarov:2006}  by adding a purely azimuthal magnetic field to build magnetised tori around rotating black holes. Dynamical evolutions of magnetized tori built with the Komissarov solution were first reported by~\cite{Montero:2007}.
On the other hand, assuming different distributions of angular momentum in the discs, \citet{Qian:2009} presented a method to build sequences of black hole thick accretion discs in dynamical equilibria, restricted however to the purely hydrodynamical case.

 In this paper we combine the two approaches considered in~\cite{Komissarov:2006} and~\citet{Qian:2009} to build new sequences of equilibrium tori around Kerr black holes. Building on these works, we present here the extension of the models of~\citet{Qian:2009} to account for discs endowed with a purely toroidal magnetic field. In our procedure we hence assume a form of the angular momentum distribution that departs from the constant case considered by~\cite{Komissarov:2006} and from which the location and morphology of the equipotential surfaces can be numerically computed. As we shall show below, for the particular case of constant angular momentum distributions, our method is in good agreement with the results of~\citet{Komissarov:2006}. We also note the recent work of~\citet{Wielgus:2015} where Komissarov's solution was extended for the particular case of power-law distributions of angular momentum. Moreover, the magnetised tori of~\citet{Wielgus:2015} were used to explore the growth of the magneto-rotational instabilty (MRI) through time-dependent numerical simulations. In particular, the long-term evolution of those tori has been recently investigated by~\citet{Fragile:2017}, who paid special attention to the decay of their magnetisation.

The organization of the paper is as follows: Section~\ref{framework} presents the analytic framework to build the discs while Section~\ref{methodology} explains the corresponding numerical procedure. The sequences of models are discussed in Section~\ref{results}. Finally, the conclusions are summarized in Section~\ref{conclusions}, where we also briefly indicate potential extensions of this work we plan to inspect. In our work we assume that the test-fluid approximation holds, thus neglecting the self-gravity of the fluid, and further assume that the spacetime is described by the Kerr metric. In mathematical expressions below Greek indices are spacetime indices running from 0 to 3 and Latin indices are only spatial. We use geometrized units where $G = c = 1$.

@@

The equations of ideal general relativistic MHD are the following conservation laws, $\nabla_{\mu} T^{\mu\nu} = 0$, $\nabla_{\mu} \,^\ast F^{\mu\nu} = 0$, and 
$\nabla_{\mu} (\rho u^{\mu}) = 0$, 
where $\nabla_{\mu}$ is the covariant derivative and
\begin{equation}\label{eq:e-m_tensor}
T^{\mu\nu} = (w + b^2)u^{\mu}u^{\nu} + \left(p + \frac{1}{2}b^2\right)g^{\mu\nu} - b^{\mu}b^{\nu},
\end{equation}
is the energy-momentum tensor of a magnetised perfect fluid, $w$ and $p$ being the fluid enthalpy density and fluid pressure, respectively. 
Moreover, $^\ast F^{\mu\nu} = b^{\mu}u^{\nu} - b^{\nu}u^{\mu}$ is the (dual of the) Faraday tensor relative to an observer with 
four-velocity $u^{\mu}$, and $b^{\mu}$ is the magnetic field in that frame, with
$b^2=b^{\mu}b_{\mu}$. Assuming the magnetic field is purely azimuthal, i.e.~$b^r = b^{\theta} = 0$,
and taking into account that the flow is stationary and axisymmetric, the conservation of the current density and of the Faraday tensor follow. Contracting Eq.~\eqref{eq:e-m_tensor} with the projection tensor $h^{\alpha}_{\,\,\beta} = \delta^{\alpha}_{\,\,\beta} + u^{\alpha}u_{\beta}$, we arrive at
\begin{equation}
(w + b^2)u_{\nu}\partial_i u^{\nu} + \partial_i\left(p + \frac{b^2}{2}\right) - b_{\nu}\partial_i b^{\nu}=0\,,
\end{equation}
where $i = r, \theta$. Following~\cite{Komissarov:2006} we rewrite this equation in terms of the specific angular momentum $l$ and of the angular velocity $\Omega$, to obtain
\begin{equation}\label{eq:diff_ver}
\partial_i(\ln u_t|) - \frac{\Omega \partial_i l}{1-l\Omega} + \frac{\partial_i p}{w} + \frac{\partial_i(\mathcal{L}b^2)}{2\mathcal{L}w} = 0\,,
\end{equation}
where $\mathcal{L} = g_{t\phi}^2 - g_{tt}g_{\phi\phi}$.
To integrate Eq.~\eqref{eq:diff_ver} we first assume a barotropic equation of state $w = w(p)$ of the form
\begin{equation}\label{eq:eos_fluid}
p = K w^{\kappa},
\end{equation}
with $K$ and $\kappa$ constants.
Then, we define the magnetic pressure as $p_{\mathrm{m}} = b^2/2$, and introduce the definitions $\tilde{p}_{\mathrm{m}} = \mathcal{L} p_{\mathrm{m}}$ and $\tilde{w} = \mathcal{L} w$, in order to write an analogue equation to Eq.~\eqref{eq:eos_fluid} for $\tilde{p}_{\mathrm{m}}$~\citep{Komissarov:2006}
\begin{equation}\label{eq:eos_mag_tilde}
\tilde{p}_{\mathrm{m}} = K_{\mathrm{m}} \tilde{w}_{\mathrm{m}}^{\lambda
},
\end{equation}
or, in terms of the magnetic pressure $p_{\mathrm{m}}$
\begin{equation}\label{eq:eos_mag}
p_{\mathrm{m}} = K_{\mathrm{m}} \mathcal{L}^{\lambda
-1} w^{\lambda
},
\end{equation}
where $K_{\mathrm{m}}$ and $\lambda
$ are constants.
This particular choice of barotropic relationships, $w = w(p)$ and $\tilde{w} = \tilde{w}(\tilde{p}_{\mathrm{m}})$, fulfill the general relativistic version of the von Zeipel theorem for a toroidal magnetic field~\citep{vonZeipel:1924, Zanotti:2015}, i.e.~the surfaces of constant $\Omega$ and constant $l$ coincide.

@@

%%%%%%%%%%%%
\section{Methodology}
\label{methodology}
%%%%%%%%%%%%

To construct our models of magnetised discs we follow the approach described in \citet{Qian:2009}. First, we find the radial location of the cusp and of the centre of the disc in the equatorial plane, $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$, defined as the solutions to the equation $l(r) - l_{\mathrm{K}} = 0$.
Next, we compute the partial derivatives of the potential, Eq.~\eqref{eq:potential}
\begin{equation}\label{eq:radial_der_pot}
\partial_r W = \partial_r \ln|u_t| - \frac{\Omega \partial_rl}{1 - \Omega l}\,,
\end{equation}
and
\begin{equation}\label{eq:polar_der_pot}
\partial_{\theta} W = \partial_{\theta} \ln|u_t| - \frac{\Omega \partial_{\theta}l}{1 - \Omega l}\,.
\end{equation}
Then, we integrate the radial partial derivative of the potential along the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ (assuming $W_{\mathrm{cusp}} = 0$) at the equatorial plane, thus obtaining the equatorial distribution of the potential between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$
\begin{equation}\label{eq:equatorial_pot}
W_{\mathrm{eq}}(r) = \int^{r_{\mathrm{c}}}_{r_{\mathrm{cusp}}}\left(\partial_r \ln|u_t| - \frac{\Omega \partial_rl}{1 - \Omega l}\right).
\end{equation}
Following~\citet{Qian:2009} (see also~\citet{Jaroszynski:1980}) we can divide equations \eqref{eq:radial_der_pot} and \eqref{eq:polar_der_pot} to obtain
\begin{equation}\label{eq:F}
F(r, \theta) = -\frac{\partial_r W}{\partial_{\theta} W} = \frac{\mathrm{d}\theta}{\mathrm{d}r}\,,
\end{equation}
where function $F(r, \theta)$ is known in closed form for the Kerr metric once an assumption for the angular momentum distribution has been made (cf.~Eq.~(\ref{eq:ansatz})). For this reason, Eq.~(\ref{eq:F}) also takes the form of an ordinary differential equation for the surfaces of constant potential, $\theta=\theta(r)$, which, upon integration, yields the location of those surfaces. Note that in~\citet{Qian:2009} these are surfaces of constant {\it fluid pressure} instead, since their discs are purely hydrodynamical.

@@

Next, we choose all the initial radial values for the integration of Eq.~\eqref{eq:F} to lie between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$ ($\theta = \pi / 2$). Since we are only interested in the equipotential surfaces inside the Roche lobe of the disc, our choice of initial values provides us a mapping of the equipotential surfaces of the torus. Given that we have already obtained both the equipotential surfaces $\theta(r)$ which cross the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ at the equatorial plane and the values of the potential in that segment, we can obtain the complete potential distribution for the torus (outside of that segment). Once we have the potential distribution, we can find the fluid pressure at the centre of the disc from Eq.~\eqref{eq:enthalpy_eq},
\begin{equation}
p_{\mathrm{c}} = w_{\mathrm{c}}(W_{\mathrm{in}} - W_{\mathrm{c}})\left(\frac{\kappa}{\kappa - 1} + \frac{\lambda
}{\lambda
 - 1}\frac{1}{\beta_{\mathrm{m}_{\mathrm{c}}}}\right)^{-1},
\end{equation}
where $w_{\mathrm{c}}$ is the enthalpy density at the centre and
\begin{equation}
\label{eq:beta_eq}
\beta_{\mathrm{m}} = p/p_{\mathrm{m}},
\end{equation}
is the magnetisation parameter ($\beta_{\mathrm{m}_{\mathrm{c}}}$ being the magnetisation parameter at the centre of the disc). Using this definition, we can find the magnetic pressure at the centre,
\begin{equation}
p_{\mathrm{m_{\mathrm{c}}}} = p_{\mathrm{c}}/\beta_{\mathrm{m}_{\mathrm{c}}}\,.
\end{equation}
With both pressures known at the centre, fluid and magnetic, we can now find the constants $K$ and $K_{\mathrm{m}}$ using equations \eqref{eq:eos_fluid} and \eqref{eq:eos_mag}. Therefore, for a given inner radius of the disc $r_{\mathrm{in}}$ we can obtain the potential $W_{\mathrm{in}}$. We now finally have all the ingredients required to find the enthalpy density distribution (Eq.~\eqref{eq:enthalpy_eq}), the fluid pressure and magnetic pressure distributions (Eqs.~\eqref{eq:eos_fluid} and \eqref{eq:eos_mag}), and the rest-mass density distribution, which can be trivially obtained inverting the barotropic relation, $p = K \rho^{\kappa}$. We note that the fluid enthalpy density $w$ includes the rest-mass density $\rho$, and can thus be defined as $w=\rho h$ where $h$ is the specific enthalpy. The relativistic definition of this quantity is $h=1+\varepsilon+p/\rho$, where $\varepsilon$ is
the specific internal energy. From a thermodynamical point of view a non-relativistic fluid satisfies $h=1$. Therefore,
since we use poytropic equations of state (relating $p$ with either $w$ or $\rho$ in the same functional form) we are
implicitly assuming that $h=1$, i.e.~the discs we build in our procedure are non-relativistic from a thermodynamical point of view.

For the integration of Eq.~\eqref{eq:equatorial_pot} we use the composite Simpson's rule. It is important to use a very small integration step because the slope is very steep. In this work, we use a step $\Delta r = 10^{-6}$. Using the analytic, constant angular momentum case for comparison, we tested that 
a larger value of $\Delta r$ gives unacceptable accuracy losses. On the other hand, to integrate the ordinary differential equation \eqref{eq:F} we use a fourth-order Runge-Kutta method. Again, it is also important here to choose a suitable step of integration, especially at the outer end of the disc, because Eq.~\eqref{eq:F} diverges at the equatorial plane (the equipotential surfaces cross the equatorial plane perpendicularly). We show the proof of this statment in Appendix~\ref{div_partial_W}.

\begin{table}
\caption{List of models for $\beta_{\mathrm{m}_{\mathrm{c}}} = 10^{3}$. From left to right the columns report: the model name (with A, B and C standing for black hole spin $a = 0.5$, $0.9$ and $0.99$, respectively), the parameters of the angular momentum distribution $\beta$ and $\gamma$, the corresponding radii of the maximum fluid pressure and magnetic pressure, $r_{{\rm{max}}}$ and $r_{{\mathrm{m, max}}}$, the gravitational potential difference between the centre and the inner radius, $\Delta W\equiv W_{\rm c}-W_{\rm in}$, and the inner and outer radii of the discs, $r_{\mathrm{in}}$ and $r_{\mathrm{out}}$.}             
\label{table:1}      
\centering          
\begin{tabular}{c c c c  c c c c}
\hline\hline       
 & $\beta$ & $\gamma$ & $r_{\rm{max}}$ &  $r_{\mathrm{m, max}}$ & $\Delta W$               & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ \\ 
 &              &                   &                          &                                        & $(\times 10^{-2})$     &                              &  \\
\hline           
A1 & $0.0$ & $0.0$ & $7.15$ &  $8.14$  & $-6.35$ & $2.91$ & -- \\ 
A2 & $0.5$ & $0.5$ & $7.15$ &  $7.65$  & $-2.27$ & $3.20$ & $11.8$\\ 
A3 & $0.9$ & $0.9$ & $7.15$ &  $7.45$  & $-0.30$ & $3.70$ &  $9.58$\\ 
A4 & $0.0$ & $0.9$ & $7.15$ &  $8.32$  & $-6.35$ & $2.91$ & $9.68$\\ 
A5 & $0.9$ & $0.0$ & $7.15$ &  $7.28$  & $-0.30$ & $3.70$ & --\\ 
 \hline 
B1 & $0.0$ & $0.0$ & $3.59$ &  $3.98$  & $-12.9$ & $1.73$ & -- \\ 
B2 & $0.5$ & $0.5$ & $3.59$ &  $3.83$  & $-4.32$ & $1.86$ & $5.35$\\ 
B3 & $0.9$ & $0.9$ & $3.59$ &  $3.73$  & $-0.54$ & $2.08$ & $4.52$\\ 
B4 & $0.0$ & $0.9$ & $3.59$ &  $3.85$  & $-12.9$ & $1.73$ & $4.54$\\ 
B5 & $0.9$ & $0.0$ & $3.59$ &  $3.81$  & $-0.54$ & $2.08$ & -- \\  
 \hline 
C1 & $0.0$ & $0.0$ & $1.98$ &  $2.28$  & $-24.6$ & $1.21$ & -- \\ 
C2 & $0.5$ & $0.5$ & $1.98$ &  $2.09$  & $-7.34$ & $1.26$ & $2.50$\\ 
C3 & $0.9$ & $0.9$ & $1.98$ &  $2.04$  & $-0.85$ & $1.35$ & $2.25$\\ 
C4 & $0.0$ & $0.9$ & $1.98$ &  $2.10$  & $-24.6$ & $1.21$ & $2.26$\\ 
C5 & $0.9$ & $0.0$ & $1.98$ &  $2.08$  & $-0.85$ & $1.35$ & --\\ 
\hline      
\end{tabular}
\end{table}

>>>>

Investigation of such systems, either by analytical or numerical means, may rely on the ability to construct suitable 
representations based on physical assumptions. The construction of equilibrium models of stationary discs around black holes has indeed a long tradition (see~\cite{Abramowicz:2013} and references therein). In particular, the `Polish doughnuts'~\citep{Abramowicz:1978,Kozlowski:1978} provide a  very general method to build equilibrium configurations of perfect fluid matter orbiting around a Kerr black hole. This fully relativistic model assumes that the disc is non-magnetised and that the matter obeys a  constant specific angular momentum distribution. This method was later extended by~\cite{Komissarov:2006}  by adding a purely azimuthal magnetic field to build magnetised tori around rotating black holes. Dynamical evolutions of magnetised tori built with the Komissarov solution were first reported by~\cite{Montero:2007}.
On the other hand, assuming different distributions of angular momentum in the discs, \citet{Qian:2009} presented a method to build sequences of black hole thick accretion discs in dynamical equilibria, restricted however to the purely hydrodynamical case.

 In this paper we have combined the two approaches considered in~\cite{Komissarov:2006} and~\citet{Qian:2009} to build new sequences of equilibrium tori around Kerr black holes. Building on these works, we present here the extension of the models of~\citet{Qian:2009} to account for discs endowed with a purely toroidal magnetic field. In our procedure we hence assumed a form of the angular momentum distribution that departs from the constant case considered by~\cite{Komissarov:2006} and from which the location and morphology of the equipotential surfaces can be numerically computed. As we shall show below, for the particular case of constant angular momentum distributions, our method is in good agreement with the results of~\citet{Komissarov:2006}. We also note the recent work of~\citet{Wielgus:2015} where Komissarov's solution was extended for the particular case of power-law distributions of angular momentum. Moreover, the magnetised tori of~\citet{Wielgus:2015} were used to explore the growth of the magneto-rotational instabilty (MRI) through time-dependent numerical simulations. In particular, the long-term evolution of those tori has been recently investigated by~\citet{Fragile:2017}, who paid special attention to the decay of their magnetisation.

The organization of the paper is as follows: Sect.~\ref{framework} presents the analytic framework to build the discs while Sect.~\ref{methodology} explains the corresponding numerical procedure. The sequences of models are discussed in Sect.~\ref{results}. Finally, the conclusions are summarised in Sect.~\ref{conclusions}, where we also briefly indicate potential extensions of this work we plan to inspect. In our work we assume that the test-fluid approximation holds, thus neglecting the self-gravity of the fluid, and further assume that the spacetime is described by the Kerr metric. In mathematical expressions below Greek indices are spacetime indices running from 0 to 3 and Latin indices are only spatial. We use geometrised units where $G = c = 1$.

@@

The equations of ideal general relativistic MHD are the following conservation laws, $\nabla_{\mu} T^{\mu\nu} = 0$, $\nabla_{\mu} \,^\ast F^{\mu\nu} = 0$, and 
$\nabla_{\mu} (\rho u^{\mu}) = 0$, 
where $\nabla_{\mu}$ is the covariant derivative and
\begin{equation}\label{eq:e-m_tensor}
T^{\mu\nu} = (w + b^2)u^{\mu}u^{\nu} + \left(p + \frac{1}{2}b^2\right)g^{\mu\nu} - b^{\mu}b^{\nu},
\end{equation}
is the energy-momentum tensor of a magnetised perfect fluid, $w$ and $p$ being the fluid enthalpy density and fluid pressure, respectively. 
Moreover, $^\ast F^{\mu\nu} = b^{\mu}u^{\nu} - b^{\nu}u^{\mu}$ is the (dual of the) Faraday tensor relative to an observer with 
four-velocity $u^{\mu}$, and $b^{\mu}$ is the magnetic field in that frame, with
$b^2=b^{\mu}b_{\mu}$. Assuming the magnetic field is purely azimuthal, that is $b^r = b^{\theta} = 0$,
and taking into account that the flow is stationary and axisymmetric, the conservation of the current density and of the Faraday tensor follow. Contracting Eq.~\eqref{eq:e-m_tensor} with the projection tensor $h^{\alpha}_{\,\,\beta} = \delta^{\alpha}_{\,\,\beta} + u^{\alpha}u_{\beta}$, we arrive at
\begin{equation}
(w + b^2)u_{\nu}\partial_i u^{\nu} + \partial_i\left(p + \frac{b^2}{2}\right) - b_{\nu}\partial_i b^{\nu}=0\,,
\end{equation}
where $i = r, \theta$. Following~\cite{Komissarov:2006} we rewrite this equation in terms of the specific angular momentum $l$ and of the angular velocity $\Omega$, to obtain
\begin{equation}\label{eq:diff_ver}
\partial_i(\ln u_t|) - \frac{\Omega \partial_i l}{1-l\Omega} + \frac{\partial_i p}{w} + \frac{\partial_i(\mathcal{L}b^2)}{2\mathcal{L}w} = 0\,,
\end{equation}
where $\mathcal{L} = g_{t\phi}^2 - g_{tt}g_{\phi\phi}$.
To integrate Eq.~\eqref{eq:diff_ver} we first assume a barotropic equation of state $w = w(p)$ of the form
\begin{equation}\label{eq:eos_fluid}
p = K w^{\kappa},
\end{equation}
with $K$ and $\kappa$ constants.
Then, we define the magnetic pressure as $p_{\mathrm{m}} = b^2/2$, and introduce the definitions $\tilde{p}_{\mathrm{m}} = \mathcal{L} p_{\mathrm{m}}$ and $\tilde{w} = \mathcal{L} w$, in order to write an analogue equation to Eq.~\eqref{eq:eos_fluid} for $\tilde{p}_{\mathrm{m}}$~\citep{Komissarov:2006}
\begin{equation}\label{eq:eos_mag_tilde}
\tilde{p}_{\mathrm{m}} = K_{\mathrm{m}} \tilde{w}_{\mathrm{m}}^{\lambda
},
\end{equation}
or, in terms of the magnetic pressure $p_{\mathrm{m}}$
\begin{equation}\label{eq:eos_mag}
p_{\mathrm{m}} = K_{\mathrm{m}} \mathcal{L}^{\lambda
-1} w^{\lambda
},
\end{equation}
where $K_{\mathrm{m}}$ and $\lambda
$ are constants.
This particular choice of barotropic relationships, $w = w(p)$ and $\tilde{w} = \tilde{w}(\tilde{p}_{\mathrm{m}})$, fulfill the general relativistic version of the von Zeipel theorem for a toroidal magnetic field~\citep{vonZeipel:1924, Zanotti:2015}, that is, the surfaces of constant $\Omega$ and constant $l$ coincide.

@@

%%%%%%%%%%%%
\section{Method}
\label{methodology}
%%%%%%%%%%%%

To construct our models of magnetised discs we have followed the approach described in \citet{Qian:2009}. First, we find the radial location of the cusp and of the centre of the disc in the equatorial plane, $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$, defined as the solutions to the equation $l(r) - l_{\mathrm{K}} = 0$.
Next, we compute the partial derivatives of the potential, Eq.~\eqref{eq:potential}
\begin{equation}\label{eq:radial_der_pot}
\partial_r W = \partial_r \ln|u_t| - \frac{\Omega \partial_rl}{1 - \Omega l}\,,
\end{equation}
and
\begin{equation}\label{eq:polar_der_pot}
\partial_{\theta} W = \partial_{\theta} \ln|u_t| - \frac{\Omega \partial_{\theta}l}{1 - \Omega l}\,.
\end{equation}
Then, we integrate the radial partial derivative of the potential along the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ (assuming $W_{\mathrm{cusp}} = 0$) at the equatorial plane, thus obtaining the equatorial distribution of the potential between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$
\begin{equation}\label{eq:equatorial_pot}
W_{\mathrm{eq}}(r) = \int^{r_{\mathrm{c}}}_{r_{\mathrm{cusp}}}\left(\partial_r \ln|u_t| - \frac{\Omega \partial_rl}{1 - \Omega l}\right).
\end{equation}
Following~\citet{Qian:2009} (see also~\citet{Jaroszynski:1980}) we can divide equations \eqref{eq:radial_der_pot} and \eqref{eq:polar_der_pot} to obtain
\begin{equation}\label{eq:F}
F(r, \theta) = -\frac{\partial_r W}{\partial_{\theta} W} = \frac{\mathrm{d}\theta}{\mathrm{d}r}\,,
\end{equation}
where function $F(r, \theta)$ is known in closed form for the Kerr metric once an assumption for the angular momentum distribution has been made (cf.~Eq.~(\ref{eq:ansatz})). For this reason, Eq.~(\ref{eq:F}) also takes the form of an ordinary differential equation for the surfaces of constant potential, $\theta=\theta(r)$, which, upon integration, yields the location of those surfaces. We note that in~\citet{Qian:2009} these are surfaces of constant {\it fluid pressure} instead, since their discs are purely hydrodynamical.

@@

Next, we choose all the initial radial values for the integration of Eq.~\eqref{eq:F} to lie between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$ ($\theta = \pi / 2$). Since we are only interested in the equipotential surfaces inside the Roche lobe of the disc, our choice of initial values provides us a mapping of the equipotential surfaces of the torus. Given that we have already obtained both the equipotential surfaces $\theta(r)$ which cross the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ at the equatorial plane and the values of the potential in that segment, we can obtain the complete potential distribution for the torus (outside of that segment). Once we have the potential distribution, we can find the fluid pressure at the centre of the disc from Eq.~\eqref{eq:enthalpy_eq},
\begin{equation}
p_{\mathrm{c}} = w_{\mathrm{c}}(W_{\mathrm{in}} - W_{\mathrm{c}})\left(\frac{\kappa}{\kappa - 1} + \frac{\lambda
}{\lambda
 - 1}\frac{1}{\beta_{\mathrm{m}_{\mathrm{c}}}}\right)^{-1},
\end{equation}
where $w_{\mathrm{c}}$ is the enthalpy density at the centre and
\begin{equation}
\label{eq:beta_eq}
\beta_{\mathrm{m}} = p/p_{\mathrm{m}},
\end{equation}
is the magnetisation parameter ($\beta_{\mathrm{m}_{\mathrm{c}}}$ being the magnetisation parameter at the centre of the disc). Using this definition, we can find the magnetic pressure at the centre,
\begin{equation}
p_{\mathrm{m_{\mathrm{c}}}} = p_{\mathrm{c}}/\beta_{\mathrm{m}_{\mathrm{c}}}\,.
\end{equation}
With both pressures known at the centre, fluid and magnetic, we can now find the constants $K$ and $K_{\mathrm{m}}$ using equations \eqref{eq:eos_fluid} and \eqref{eq:eos_mag}. Therefore, for a given inner radius of the disc $r_{\mathrm{in}}$ we can obtain the potential $W_{\mathrm{in}}$. We now finally have all the ingredients required to find the enthalpy density distribution (Eq.~\eqref{eq:enthalpy_eq}), the fluid pressure and magnetic pressure distributions (Eqs.~\eqref{eq:eos_fluid} and \eqref{eq:eos_mag}), and the rest-mass density distribution, which can be trivially obtained inverting the barotropic relation, $p = K \rho^{\kappa}$. We note that the fluid enthalpy density $w$ includes the rest-mass density $\rho$, and can thus be defined as $w=\rho h$ where $h$ is the specific enthalpy. The relativistic definition of this quantity is $h=1+\varepsilon+p/\rho$, where $\varepsilon$ is
the specific internal energy. From a thermodynamical point of view a non-relativistic fluid satisfies $h=1$. Therefore,
since we use polytropic equations of state (relating $p$ with either $w$ or $\rho$ in the same functional form) we are
implicitly assuming that $h=1$, that is, the discs we build in our procedure are non-relativistic from a thermodynamical point of view.

For the integration of Eq.~\eqref{eq:equatorial_pot} we have used the composite Simpson's rule. It is important to use a very small integration step because the slope is very steep. In this work, we have used a step $\Delta r = 10^{-6}$. Using the analytic, constant angular momentum case for comparison, we tested that 
a larger value of $\Delta r$ gives unacceptable accuracy losses. On the other hand, to integrate the ordinary differential equation \eqref{eq:F} we have employed a fourth-order Runge-Kutta method. Again, it is also important here to choose a suitable step of integration, especially at the outer end of the disc, because Eq.~\eqref{eq:F} diverges at the equatorial plane (the equipotential surfaces cross the equatorial plane perpendicularly). We show the proof of this statment in Appendix~\ref{div_partial_W}.

\begin{table}
\caption{Models for $\beta_{\mathrm{m}_{\mathrm{c}}} = 10^{3}$. From left to right the columns report: the model name (with A, B and C standing for black hole spin $a = 0.5$, $0.9$ and $0.99$, respectively), the parameters of the angular momentum distribution $\beta$ and $\gamma$, the corresponding radii of the maximum fluid pressure and magnetic pressure, $r_{{\rm{max}}}$ and $r_{{\mathrm{m, max}}}$, the gravitational potential difference between the centre and the inner radius, $\Delta W\equiv W_{\rm c}-W_{\rm in}$, and the inner and outer radii of the discs, $r_{\mathrm{in}}$ and $r_{\mathrm{out}}$.}             
\label{table:1}      
\centering          
\begin{tabular}{c c c c  c c c c}
\hline\hline       
 & $\beta$ & $\gamma$ & $r_{\rm{max}}$ &  $r_{\mathrm{m, max}}$ & $\Delta W$               & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ \\ 
 &              &                   &                          &                                        & $(\times 10^{-2})$     &                              &  \\
\hline           
A1 & $0.0$ & $0.0$ & $7.15$ &  $8.14$  & $-6.35$ & $2.91$ & -- \\ 
A2 & $0.5$ & $0.5$ & $7.15$ &  $7.65$  & $-2.27$ & $3.20$ & $11.8$\\ 
A3 & $0.9$ & $0.9$ & $7.15$ &  $7.45$  & $-0.30$ & $3.70$ &  $9.58$\\ 
A4 & $0.0$ & $0.9$ & $7.15$ &  $8.32$  & $-6.35$ & $2.91$ & $9.68$\\ 
A5 & $0.9$ & $0.0$ & $7.15$ &  $7.28$  & $-0.30$ & $3.70$ & --\\ 
 \hline 
B1 & $0.0$ & $0.0$ & $3.59$ &  $3.98$  & $-12.9$ & $1.73$ & -- \\ 
B2 & $0.5$ & $0.5$ & $3.59$ &  $3.83$  & $-4.32$ & $1.86$ & $5.35$\\ 
B3 & $0.9$ & $0.9$ & $3.59$ &  $3.73$  & $-0.54$ & $2.08$ & $4.52$\\ 
B4 & $0.0$ & $0.9$ & $3.59$ &  $3.85$  & $-12.9$ & $1.73$ & $4.54$\\ 
B5 & $0.9$ & $0.0$ & $3.59$ &  $3.81$  & $-0.54$ & $2.08$ & -- \\  
 \hline 
C1 & $0.0$ & $0.0$ & $1.98$ &  $2.28$  & $-24.6$ & $1.21$ & -- \\ 
C2 & $0.5$ & $0.5$ & $1.98$ &  $2.09$  & $-7.34$ & $1.26$ & $2.50$\\ 
C3 & $0.9$ & $0.9$ & $1.98$ &  $2.04$  & $-0.85$ & $1.35$ & $2.25$\\ 
C4 & $0.0$ & $0.9$ & $1.98$ &  $2.10$  & $-24.6$ & $1.21$ & $2.26$\\ 
C5 & $0.9$ & $0.0$ & $1.98$ &  $2.08$  & $-0.85$ & $1.35$ & --\\ 
\hline      
\end{tabular}
\end{table}

\begin{table}
\caption{Models for $\beta_{\mathrm{m}_{\mathrm{c}}} = 1$. The naming of the models and the parameters reported in the columns are as in Table~\ref{table:1}.}             
\label{table:2}      
\centering          
\begin{tabular}{c c c c  c c c c}
\hline\hline       
 & $\beta$ & $\gamma$ & $r_{\rm{max}}$ &  $r_{\mathrm{m, max}}$ & $\Delta W$               & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ \\ 
 &              &                   &                          &                                        & $(\times 10^{-2})$     &                              &  \\
\hline           
A1 & $0.0$ & $0.0$ & $6.00$ &  $6.49$  & $-6.35$ & $2.91$ & -- \\ 
A2 & $0.5$ & $0.5$ & $6.25$ &  $6.66$  & $-2.27$ & $3.20$ & $11.8$\\ 
A3 & $0.9$ & $0.9$ & $6.50$ &  $6.80$  & $-0.30$ & $3.70$ &  $9.58$\\ 
A4 & $0.0$ & $0.9$ & $6.00$ &  $6.49$  & $-6.35$ & $2.91$ & $9.68$\\ 
A5 & $0.9$ & $0.0$ & $6.50$ &  $6.80$  & $-0.30$ & $3.70$ & --\\ 
 \hline 
B1 & $0.0$ & $0.0$ & $3.02$ &  $3.26$  & $-12.9$ & $1.73$ & -- \\ 
B2 & $0.5$ & $0.5$ & $3.16$ &  $3.35$  & $-4.32$ & $1.86$ & $5.35$\\
B3 & $0.9$ & $0.9$ & $3.29$ &  $3.43$  & $-0.54$ & $2.08$ & $4.52$\\ 
B4 & $0.0$ & $0.9$ & $3.02$ &  $3.26$  & $-12.9$ & $1.73$ & $4.54$\\  
B5 & $0.9$ & $0.0$ & $3.29$ &  $3.43$  & $-0.54$ & $2.08$ & -- \\ 
 \hline 
C1 & $0.0$ & $0.0$ & $1.68$ &  $1.80$  & $-24.6$ & $1.21$ & -- \\ 
C2 & $0.5$ & $0.5$ & $1.77$ &  $1.86$  & $-7.34$ & $1.26$ & $2.50$\\  
C3 & $0.9$ & $0.9$ & $1.84$ &  $1.91$  & $-0.85$ & $1.35$ & $2.25$\\ 
C4 & $0.0$ & $0.9$ & $1.68$ &  $1.80$  & $-24.6$ & $1.21$ & $2.26$\\ 
C5 & $0.9$ & $0.0$ & $1.84$ &  $1.91$  & $-0.85$ & $1.35$ & --\\ 
\hline      
\end{tabular}
\end{table}
